Clique and Constraint Models for Maximum Common (Connected) Subgraph Problems

### 2016-03-24 paper/paper.tex
Pictures and text

\usepackage{microtype}
\usepackage{complexity}                % \P, \NP etc
\usepackage{tikz}                      % For pretty pictures
\usepackage{amsmath}                   % \operatorname
\usepackage{hyperref}                  % clicky links
\usepackage{cleveref}                  % no need to type Figure etc
\usepackage{gnuplot-lua-tikz}          % graphs
\usepackage[ruled,vlined]{algorithm2e} % algorithms (after cleverref!)
\usepackage{placeins}                  % \FloatBarrier
\usepackage{cite}

@@

Max common subgraph ?? cite some history and why we care

Connected \cite{DBLP:journals/tcs/Koch01,DBLP:conf/mco/VismaraV08}

CP models \cite{DBLP:conf/cp/NdiayeS11}

Clique models, discuss below.

@@

We instead use a maximum clique algorithm. There has been a lot of recent
research on this
\cite{DBLP:conf/dmtcs/TomitaS03,DBLP:journals/jgo/TomitaK07,DBLP:conf/walcom/TomitaSHTW10,DBLP:journals/cor/SegundoRJ11,DBLP:journals/algorithms/Prosser12,DBLP:journals/ol/SegundoMRH13,DBLP:conf/ictai/LiFX13,DBLP:journals/cor/SegundoT14,DBLP:conf/lion/SegundoLB14,DBLP:conf/cp/McCreeshP14,DBLP:journals/jco/BatsynGMP14,DBLP:journals/cor/SegundoNB15,DBLP:conf/lion/NikolaevBS15,DBLP:conf/lion/LiJX15}.

>>>>

\usepackage{microtype}
\usepackage{complexity}                % \P, \NP etc
\usepackage{tikz}                      % For pretty pictures
\usepackage{amsmath}                   % \operatorname
\usepackage{hyperref}                  % clicky links
\usepackage{cleveref}                  % no need to type Figure etc
\usepackage{gnuplot-lua-tikz}          % graphs
\usepackage[ruled,vlined]{algorithm2e} % algorithms (after cleverref!)
\usepackage{placeins}                  % \FloatBarrier
\usepackage{cite}
\usepackage[bottom]{footmisc}

@@

\begin{figure}[b] % first page if possible
    \centering
    \begin{tikzpicture}[scale=0.4]%{{{
        \definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
        \definecolor{uofgsunshine}{rgb}{1.0, 0.862745, 0.211765}
        \definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
        \begin{scope}
            \node[draw, circle, fill=uofgsunshine, inner sep=1pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=1pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=1pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=1pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=1pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!40, fill=white, inner sep=1pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=1pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=1pt, font=\normalsize] (M8) at ($(330:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=1pt, font=\normalsize] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};

            \draw [very thick] (M1) -- (M2);
            \draw [very thick] (M2) -- (M4);
            \draw [very thick] (M3) -- (M5);
            \draw [color=uofgsandstone!40] (M4) -- (M6);
            \draw [color=uofgsandstone!40] (M5) -- (M6);
            \draw [very thick] (M3) -- (M1);
            \draw [color=uofgsandstone!40] (M6) -- (M7);
            \draw [color=uofgsandstone!40] (M6) -- (M8);
            \draw [color=uofgsandstone!40] (M6) -- (M9);
            \draw [very thick] (M7) -- (M9);
            \draw [very thick] (M8) -- (M9);
        \end{scope}

        \begin{scope}[xshift=5cm]
            \node[draw, circle, fill=uofgsunshine, inner sep=1pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=1pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=1pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=1pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=1pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!40, fill=white, inner sep=1pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=1pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=1pt, font=\normalsize] (M8) at ($(270:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=1pt, font=\normalsize] (M9) at ($(330:1.5) + (M8)$) {\phantom{0}};

            \draw [very thick] (M1) -- (M2);
            \draw [very thick] (M2) -- (M4);
            \draw [very thick] (M3) -- (M5);
            \draw [color=uofgsandstone!40] (M4) -- (M6);
            \draw [color=uofgsandstone!40] (M5) -- (M6);
            \draw [very thick] (M3) -- (M1);
            \draw [color=uofgsandstone!40] (M6) -- (M7);
            \draw [very thick] (M7) -- (M8);
            \draw [very thick] (M8) -- (M9);
        \end{scope}

        \begin{scope}[xshift=12cm]
            \node[draw, circle, fill=uofgcobalt, inner sep=1pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=1pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=1pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=1pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=1pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=1pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!40, fill=white, inner sep=1pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=1pt, font=\normalsize] (M8) at ($(330:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!40, fill=white, inner sep=1pt, font=\normalsize] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};

            \draw [very thick] (M1) -- (M2);
            \draw [very thick] (M2) -- (M4);
            \draw [very thick] (M3) -- (M5);
            \draw [very thick] (M4) -- (M6);
            \draw [very thick] (M5) -- (M6);
            \draw [very thick] (M3) -- (M1);
            \draw [color=uofgsandstone!40] (M6) -- (M7);
            \draw [very thick] (M6) -- (M8);
            \draw [color=uofgsandstone!40] (M6) -- (M9);
            \draw [color=uofgsandstone!40] (M7) -- (M9);
            \draw [color=uofgsandstone!40] (M8) -- (M9);
        \end{scope}

        \begin{scope}[xshift=17cm]
            \node[draw, circle, fill=uofgcobalt, inner sep=1pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=1pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=1pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=1pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=1pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=1pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=1pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!40, fill=white, inner sep=1pt, font=\normalsize] (M8) at ($(270:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!40, fill=white, inner sep=1pt, font=\normalsize] (M9) at ($(330:1.5) + (M8)$) {\phantom{0}};

            \draw [very thick] (M1) -- (M2);
            \draw [very thick] (M2) -- (M4);
            \draw [very thick] (M3) -- (M5);
            \draw [very thick] (M4) -- (M6);
            \draw [very thick] (M5) -- (M6);
            \draw [very thick] (M3) -- (M1);
            \draw [very thick] (M6) -- (M7);
            \draw [draw=uofgsandstone!40] (M7) -- (M8);
            \draw [draw=uofgsandstone!40] (M8) -- (M9);
        \end{scope}
\end{tikzpicture}
\caption{A maximum common induced subgraph of the first two graphs has eight vertices. There are
four symmetric solutions, all involving the shaded vertices. However, if we require the common
subgraph to be connected, only seven vertices may be selected---one way to do this is shown in the
third and fourth graphs.}\label{figure:mcsexample}
\end{figure}

A common subgraph of two graphs is a graph which is subgraph-isomorphic to both graphs---we
illustrate this concept in \cref{figure:mcsexample}, and define this formally below. Finding a
maximum common subgraph gives a measure of the similarity between two graphs, and allows for the
edit distance to be computed ?? cite. Maximum common subgraph problems have applications in
chemistry, computer vision, ?? cite.

One approach to this problem is via constraint programming (CP) models
\cite{DBLP:conf/cp/NdiayeS11}. These models are based upon constructing an injective mapping from
the vertices of one graph (typically the smaller) to the other, in a similar manner to subgraph
isomorphism algorithms ?? cite. However, to allow for a partial mapping, a special null value is
added to each domain (and the injectivity condition does not apply to null values). The problem then
becomes an optimisation problem, where we seek to minimise the number of null values. ?? find a way
to work weak all diff in here. Unfortunately, this null value invalidates many of the invariants
used for subgraph isomorphism---we may no longer reason using the degree of vertices, or the
distances between them, for example, and so the problem can be much more challenging in practice.

Another popular approach is to reduce the problem to finding a maximum clique (or sometimes,
equivalently, an independent set or a vertex cover) in an association graph (which we describe
below). This seems to be less effective in practice, but previous experimental evaluations have used
weak clique algorithms, or even clique enumeration algorithms. Maximum clique algorithms are an
active research area
\cite{DBLP:conf/dmtcs/TomitaS03,DBLP:journals/jgo/TomitaK07,DBLP:conf/walcom/TomitaSHTW10,DBLP:journals/cor/SegundoRJ11,DBLP:journals/algorithms/Prosser12,DBLP:journals/ol/SegundoMRH13,DBLP:conf/ictai/LiFX13,DBLP:journals/cor/SegundoT14,DBLP:conf/lion/SegundoLB14,DBLP:conf/cp/McCreeshP14,DBLP:journals/jco/BatsynGMP14,DBLP:journals/cor/SegundoNB15,DBLP:conf/lion/NikolaevBS15,DBLP:conf/lion/LiJX15,DBLP:journals/jcc/KoncDTRJ12,DBLP:journals/algorithms/McCreeshP13,DBLP:journals/topc/McCreeshP15,DBLP:journals/cor/SegundoLP16},
and so the first half of this paper re-evaluates this approach using a modern algorithm.

(Other approaches have been tried, including MIP ?? cite, SAT.)

A particularly challenging variant of the maximum common subgraph problem requires the selected
subgraph to be connected \cite{DBLP:journals/tcs/Koch01,DBLP:conf/mco/VismaraV08}. The second half
of the paper looks at this variant. In principle, for a CP approach, we may simply add a ``forms a
connected graph'' global constraint to the model. However, we will show that better results may be
obtained via a special branching rule. We then explain how to adapt this branching rule to run
inside a clique-like algorithm, to solve the problem via an association graph.

@@

We instead use a maximum clique algorithm. There has been a lot of recent
research on this
.
